---
title: "Exercise1"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("sf")
#install.packages("raster")
#install.packages("tmap")
#install.packages("spData")
#install.packages("spDataLarge", repos = "https://nowosad.github.io/drat/", type = "source")

library(sf)
library(raster)
library(tmap)
library(spData)
library(spDataLarge)
```

## Existing code to be edited
```{r}
tm_shape(nz_elev)  +
  tm_raster(title = "elev", 
            style = "cont",
            palette = "BuGn") +
  tm_shape(nz) +
  tm_borders(col = "red", 
             lwd = 3) +
  tm_scale_bar(breaks = c(0, 100, 200),
               text.size = 1) +
  tm_compass(position = c("LEFT", "center"),
             type = "rose", 
             size = 2) +
  tm_credits(text = "A. Sobotkova, 2020") +
  tm_layout(main.title = "My map",
            bg.color = "orange",
            inner.margins = c(0, 0, 0, 0))
```

## Exercise I

1. Change the map title from "My map" to "New Zealand".
2. Update the map credits with your own name and today's date.
3. Change the color palette to "-RdYlGn". 
   (You can also try other palettes from http://colorbrewer2.org/)
4. Put the north arrow in the top right corner of the map.
5. Improve the legend title by adding the used units (m asl).
6. Increase the number of breaks in the scale bar.
7. Change the borders' color of the New Zealand's regions to black. 
   Decrease the line width.
8. Change the background color to any color of your choice.

```{r}
tm_shape(nz_elev)  +
  tm_raster(title = "Elev. meters asl", 
            style = "cont",
            palette = "-RdYlGn") +
  tm_shape(nz) +
  tm_borders(col = "black", 
             lwd = 1) +
  tm_scale_bar(breaks = c(0, 50, 100, 150, 200),
               text.size = 1) +
  tm_compass(position = c("RIGHT", "top"),
             type = "rose", 
             size = 2) +
  tm_credits(text = "M.B.B, 2021-02-11") +
  tm_layout(main.title = "New Zealand",
            bg.color = "hotpink",
            inner.margins = c(0, 0, 0, 0))
```

## Exercise II

9. Read two new datasets, `srtm` and `zion`, using the code below.
   To create a new map representing these datasets.

```{r}
srtm = raster(system.file("raster/srtm.tif", package = "spDataLarge"))
zion = read_sf(system.file("vector/zion.gpkg", package = "spDataLarge"))

tm_shape(srtm)  +
  tm_raster(title = "Elev. m. asl", 
            style = "cont",
            palette = "-RdYlGn") +
  tm_shape(zion) +
  tm_borders(col = "black", 
             lwd = 1) +
  tm_scale_bar(breaks = c(0, 2, 4, 6, 8),
  						 position = c("LEFT", "bottom"),
               text.size = 1) +
  tm_compass(position = c("RIGHT", "top"),
             type = "rose", 
             size = 2) +
  tm_credits(text = "M.B.B, 2021-02-11",
  					 position = c("LEFT", "bottom"),) +
  tm_layout(main.title = "Zion National Park",
  					legend.position = c("left","center"),
            bg.color = "hotpink",
            inner.margins = c(0, 0, 0, 0))
```











